--- 
site: bookdown::bookdown_site
documentclass: book
classoption: oneside
#bibliography: [book.bib, pakages.bib]
biblio-style: apalike
link-citations: yes
colorlinks: yes
lot: yes
lof: yes
fontsize: 11pt
#github-repo: tbd/tbd
---

```{r, include=FALSE}
library(readxl)
library(knitr)
library(kableExtra)
library(readr)
library(ggplot2)
library(dplyr)
library(scales)
library(glue)
```

# Solvents: Asphalt {#asphalt}

## Sector Descriptions and Overview 

Liquid asphalt is a petroleum-derived substance used in paving applications, such as the construction of roads, parking lots, driveways, and airport runways, as well as non-paving applications, such as the manufacturing of roofing shingles. While liquid asphalt can be found in natural deposits, most is produced from crude oil. Vacuum distillation separates components of crude oil based on boiling point. Products generated from this process include naptha, gasoline, diesel, and liquid asphalt, the last of which has a boiling point greater than 500 °C. As a result, most volatile, light fractions of organics are separated from liquid asphalt during distillation and prior to use. 

In paving applications, liquid asphalt can be applied cold or heated. If applied cold, additional components must be added to lower the viscosity of the material, which allows it to be spread upon a surface (e.g., roadway surface). Cutback asphalt (SCC: 2461021000) is a cold application process that involves mixing the liquid asphalt with petroleum solvents (e.g., naphtha, kerosene, fuel oil, diesel, etc.). Following application, these higher volatility solvents evaporate, leaving the asphalt in place. Due to this increased organic emissions potential, cutback asphalts have grown less common over time and this process now constitutes ~1% of liquid asphalt is use [[ref 1]](#asphalt-references). Emulsified asphalt (SCC: 2461022000) is a separate cold application process that utilizes water-based solvents and an emulsifying agent. The result is a stable liquid suspension with asphalt globules. Following application, the additives evaporate and leave the asphalt in place. In contrast to cutback asphalts, emulsified asphalts have become more common in recent years and ~10% of liquid asphalt is now used in these applications.

Liquid asphalt can also be applied heated, which both lowers the viscosity of the material and minimizes the need for added solvents. Hot-mix asphalt application (SCC: 2461025100), which is the traditional method for asphalt pavement production, involves combining the liquid asphalt with aggregate at a hot-mix plant and heating the mixture to +150 °C. The mixture is then hauled to the usage site heated, where it is placed, compacted, and ambiently cooled. This process does not require any solvent additions. Warm-mix asphalt application (SCC: 2461025200) is a more recent technology that enables asphalt pavement production to occur at 20 – 40 °C cooler temperatures than hot-mix asphalt application. Warm-mix asphalt applications at reduced temperatures now constitute ~18% of asphalt paving applications [[ref 2]](#asphalt-references). These lower production temperatures promote energy savings through reductions in fuel use and lower emissions at the hot-mix plant. To reduce the viscosity of the liquid asphalt in warm-mix applications, water, water-bearing minerals, chemicals, waxes, organic additives, or a combination of technologies must be added [[ref 3]](#asphalt-references). 

While heated applications processes represent most liquid asphalt paving applications, it has historically been assumed that emissions were sparse due to the removal of the more volatile organics during distillation. Recent research has demonstrated that less volatile organic vapors from liquid asphalt do evaporate at temperatures associated with hot-mix application (~140 °C), warm-mix application (120 °C), and post-application, or “in-use” temperatures [[ref 4]](#asphalt-references). Emission during the “in-use” period diffuse from the pavement over time following application. 

Roofing asphalts include asphalt cements and emulsions used in the manufacturing of asphalt shingles, asphalt sealant, and roof tar. In 2020, manufacturing of these products consumed ~1.95 million short tons of asphalt, which is ~9% of all asphalt generated in the United States. These materials are predominantly applied in ambient conditions (i.e., as asphalt shingles). Roof tar, which is applied hot and at temperatures like hot-mix road asphalt (+150 °C), comprises ~5% of roofing asphalt usage. In addition, all forms of roofing asphalt are exposed to both high temperatures and solar radiation throughout the duration of their life cycles, which have the potential of generating enhancements of emissions.

Table \@ref(tab:asphalt-sccs) notes all SCCs covered in this source category and the SCCs for which the EPA generates default emissions. 

```{r, include=FALSE}
df <- read_excel("./tables/Section-31/asphalt-sccs.xlsx")
df[is.na(df)] <- ""
```

```{r tidy=FALSE, echo=FALSE, booktabs = TRUE}
kableExtra::kbl(
  df, booktabs = TRUE,
  caption = 'Asphalt SCCs in the 2023 NEI.',
  label = 'asphalt-sccs'
) %>%
kable_styling(full_width = FALSE, position = "center", latex_options = "scale_down")
```

## EPA-developed estimates

Usage of liquid asphalt at the state-level for each process is calculated and subsequently allocated to the county-level using estimated vehicle miles traveled on paved roads and construction expenditure statistics for roofing processes. Emission factors consider both application and in-use processes. Net county-level emissions are quantified by multiplying the SCC-specific liquid asphalt usage by SCC-specific emission factors. The sources of data, calculation of state-level, SCC-specific usage results, allocation of state-level usage to the county-level, emission factors, and emission estimates are all discussed in subsequent sections.

### Activity Data

Activity data for these sources are the amount of liquid asphalt used in each process. This includes cutback, emulsified, hot mix, warm mix, and roofing. Each year, the Asphalt Institute releases an asphalt usage survey for the Unites States and Canada that reflects usage among their membership [[ref 1]](#asphalt-references). The Asphalt Institute estimates that their membership captures 90% of the United States market and the survey includes usage for cutback, emulsified, a summation of heated application processes, and roofing asphalt at the level of Petroleum Administration for Defense Districts (PADD) and sub-PADD levels. PADDs are geographic aggregations of the 50 states and the District of Columbia that were generated during World War II for purposes of administering oil allocation. At the sub-PADD level, usage resolution is provided in aggregations that include up to six states. Following 2020, the Asphalt Institute no longer publicly released this annual survey. As such, additional methods were used to scale the 2020 Asphalt Institute survey to 2023 values, which are further described below.

Separately, state-level, year-specific data on the production of asphalt pavement (i.e., mass of asphalt binder and aggregate) for heated application and the proportions used in warm mix processes at reduced temperatures are available from a National Asphalt Pavement Association (NAPA) annual report [[ref 2]](#asphalt-references). This report is not used in-lieu of the Asphalt Institute report because it does not include usage of cutback, emulsified, and roofing asphalts. However, the state-level heated application usage and reduced temperature (i.e., warm mix usage) proportions are used to allocate cutback, emulsified, hot mix, and warm mix usage from the Asphalt Institute survey to the state-level. Therefore, it is assumed that the state-level proportions of liquid paving asphalt used in heated applications within a sub-PADD match the state-level proportions of liquid asphalt used in cold application (i.e., cutback and emulsified) within a sub-PADD. Furthermore, since 2023 data is available from the NAPA annual report, the state-level growth or decline in asphalt pavement usage between 2020 and 2023 from these reports are used to scale the 2020 Asphalt Institute survey to generate 2023 estimates of all paving processes. Allocation of activity data for roofing asphalt utilizes separate methods and is described below. The derivation of state-level, per-application usage is as follows:

\begin{equation} 
  Usage_{s,a} = Usage_{sp,a} \times \frac{HA_{s}}{HA_{sp}}
  (\#eq:state-usage)
\end{equation} 

<div style="margin-left: 2em;">
  Where: <br> \newline
  \( Usage_{s,a} \) = Liquid asphalt usage in state *s* for application *a*, in short tons <br> \newline
  \( Usage_{sp,a} \) = sub-PADD usage of liquid asphalt in sub-PADD *sp* associated with state *s* for application *a* from the Asphalt Institute survey, in short tons <br> \newline
  \( HA_{s} \) = Heated application usage in state *s* from the NAPA survey, in short tons <br> \newline
  \( HA_{sp} \) = Heated application usage in sub-PADD *sp* associated with state *s* from the NAPA survey, in short tons <br> \newline
  \( a \) = Application types include hot-mix, warm-mix, cutback, and emulsified <br> \newline
  \( sp \) = sub-PADD associated with state *s*. sub-PADDs include the 11 districts included in the Asphalt Institute survey [[ref 1]](#ref-1) <br> \newline
</div>

An additional transformation must be done to split the state-level, heated application usage into hot-mix and warm-mix application, respectively.

\begin{equation} 
  Usage_{s,a} = 
  \begin{aligned}
  &\begin{cases} 
    Usage_{s} \times \frac{WMA_{s}}{HA_{s}} & \text{for warm-mix application}\\ 
    Usage_{s} \times \frac{(1 - WMA_{s})}{HA_{s}} & \text{for hot-mix application }
  \end{cases}
  \end{aligned}
  (\#eq:piecewise-usage)
\end{equation}

<div style="margin-left: 2em;">
  Where: <br> \newline
  \( Usage_{s,a} \) = Liquid asphalt usage in state *s* for application *a*, in short tons <br> \newline
  \( Usage_{s} \) = Liquid asphalt usage in state *s* for heated application, as derived in \@ref(eq:state-usage), in short tons <br> \newline
  \( WMA_{s} \) = Warm-mix application usage at reduced temperatures in state *s* from the NAPA survey, in short tons <br> \newline
  \( HA_{s} \) = Heated application usage in state *s* from the NAPA survey, in short tons. <br> \newline
  \( a \) = Application types include hot-mix and warm-mix. State-level summation across application types in Eqn. \@ref(eq:piecewise-usage) yield the state-level application usage for heated application from Eqn. \@ref(eq:state-usage). <br> \newline
</div>

### Allocation Procedure

State-level asphalt usage for paving applications is allocated to the county-level using vehicular miles traveled on paved roads. As such, it is assumed that county-level paving activity is proportional to the estimated vehicular miles traveled on paved roads within each county in a state. Here, proprietary telematics data was used to generate county-specific, VMT totals on various road-types. Road-types with descriptions associated with paved/unpaved roads were separately summed, with the residual (“unknown”) VMT proportioned to the paved/unpaved assortments using the prior totals. Therefore, it is assumed that the “unknown” VMT is proportional to the “known” paved/unpaved VMT totals. 

County-level paved VMT estimates within a state are then paired with the usage estimates from Eqn. \@ref(eq:state-usage) - \@ref(eq:piecewise-usage) to yield county-level asphalt usage for each process (hot mix, warm mix, cutback, and emulsified) as follows:

\begin{equation} 
  Usage_{c,a} = Usage_{s,a} \times \frac{PVMT_{c}}{PVMT_{s}}
  (\#eq:county-usage)
\end{equation} 

<div style="margin-left: 2em;">
  Where: <br> \newline
  \( Usage_{c,a} \) = Liquid asphalt usage in county *c* for application *a*, in short tons. <br> \newline
  \( Usage_{s,a} \) = Liquid asphalt usage in state *s* associated with county *c* for application *a*, in short tons (from Eqn. \@ref(eq:state-usage) - \@ref(eq:piecewise-usage)) <br> \newline
  \( PVMT_{c} \) = Estimated paved vehicular miles traveled in county *c*, in miles <br> \newline
  \( PVMT_{s} \) = Estimated paved vehicular miles traveled in state *s* associated with county *c*, in miles <br> \newline
  \( a \) = Application types include hot-mix, warm-mix, cutback, and emulsified <br> \newline
</div>

Activity data for roofing asphalt is retrieved from the Asphalt Institute report at the sub-PADD level and allocated to the state-level using nonresidential [[ref 6]](#asphalt-references) and state/local construction-put-in-place [[ref 7]](#asphalt-references) statistics from the U.S. Census Bureau’s Construction Spending datasets. The state-level 2023 and 2020 expenditure values from these datasets are used to scale the 2020 Asphalt Institute roofing asphalt usage numbers to 2023 estimates. Then, state-level roofing asphalt usage is allocated to the county-level using population data from the U.S. Census Bureau [[ref 8]](#asphalt-references) Roofing asphalt usage spans multiple material-types (e.g., asphalt shingles and roof tar) and it is assumed that 5% of roofing asphalt is used for roof tar, with the remaining 95% predominantly being used for shingles. These roof tar/roofing shingles proportions are included in the final weighted emissions factors for 2461023000.

### Emission Factors

Emission factors for all paving processes (hot-mix, warm-mix, cutback, and emulsified) captures emissions that occur during application and in-use. Both application and in-use emission factors for hot-mix and warm-mix asphalt, as well as the in-use emission factors for cutback and emulsified asphalt, were retrieved from Khare et al., 2020 [[ref 4]](#asphalt-references). Emission factors associated with application for cutback and emulsified asphalt paving were not updated. 

During the hot-mix application process, the asphalt pavement (i.e., mixture of liquid asphalt and aggregate) is heated and applied at elevated temperatures (~150 °C). Emissions are highest when sustained heating is initiated and exponentially decline thereafter. Measurements indicate that the exponential function below (Eqn. \@ref(eq:hot-application)) fits the dynamic change in emissions over a prolonged experiment (> 6 days). However, hot-mix asphalt is not heated for prolonged periods. Here, it is assumed that the application process takes 5 hours and is meant to capture the time between transport, paving, and ambient cooling. 

\begin{equation} 
  EF = 7.7 \times \exp^{-0.016t} + 16 \times \exp^{-0.5t}
  (\#eq:hot-application)
\end{equation} 

<div style="margin-left: 2em;">
  Where: <br> \newline
  \( EF \) = Emission factor of gas-phase organics, in mg per min per kg asphalt <br> \newline
  \( t \) = Time, in hours <br> \newline
</div>

Integrating Eqn. \@ref(eq:hot-application) over a period of 5 hours yields an emission factor of 4 g/kg asphalt, or 8.04 lb/short ton asphalt (i.e., 0.4% emissions by weight), and represents the emissions from hot-mix asphalt during the application process. The warm-mix asphalt application process generally occurs at 20 – 40 °C cooler temperatures than hot-mix asphalt application. Reducing the asphalt temperature from 140 °C to 120 °C reduced the initial pulse of emissions by ~46% (Fig. S5 of [ref 4](#asphalt-references)). As such, a warm-mix application emission factor of 2 g/kg asphalt, or 4.32 lb/short ton asphalt (i.e., 0.2% emissions by weight), is adopted.

Cutback and emulsified application emission factors are developed using compositional information from material safety and data sheets (MSDS) for cutback [[ref 9]](#asphalt-references) and emulsified [[ref 10]](#asphalt-references) asphalt. Assuming a volatilization fraction of 95% for all components yields an emission factor of 813.96 lb/short ton asphalt (407 g/kg asphalt) for cutback applications and 195.51 lb/short ton asphalt (98 g/kg asphalt) for emulsified applications.

In-use emissions follow application and occur under ambient temperatures. Since emissions are strongly influenced by temperature, climatological variation can impact the speed in which emissions occur. Measurements associated with a sustained heating experiment at 60 °C feature an exponential decline and fit the function below (Eqn. \@ref(eq:hot-in-use)). While 60 °C is above ambient conditions for all locations within the United States, measurements show that emissions flatten within a day and remain near-constant for more than 2 additional days. Here, it is assumed that emissions within 72-hours under 60 °C will occur within 1-year under ambient conditions at all locations within the United States. Integrating Equation 5 over a period of 72 hours yields an emission factor of 1 g/kg asphalt, or 2.01 lb/short ton asphalt (i.e., 0.1% emissions by weight), and represents the emissions from all in-use asphalt paving process.

\begin{equation} 
  EF = 0.1 + 3.3 \times \exp^{-0.35t}
  (\#eq:hot-in-use)
\end{equation} 

<div style="margin-left: 2em;">
  Where: <br> \newline
  \( EF \) = Emission factor of gas-phase organics, in mg per min per kg asphalt <br> \newline
  \( t \) = Time, in hours <br> \newline
</div>

Taken together, the VOC emission factors for all asphalt paving process are the summation of emissions associated with application and in-use. 

Emission factors for roofing asphalt materials are retrieved from Khare et al., 2020 [[ref 4]](#asphalt-references). Emissions associated with application for roof tar and in-use emissions from asphalt shingles and roof tar are separately considered and summed for total roofing asphalt emissions. Roof tar is applied at similar temperatures as hot-mix asphalt used in paving applications. Therefore, we assume a similar application emissions factor (4.0 g / kg asphalt; 8.0 lb / ton asphalt). For in-use emissions, we use the emission factors for asphalt shingles and roofing asphalt under “no sun” conditions from Khare et al., 2020 (Table S7) and apply the duration of the experiment (46 hours) to the reported emissions factor (5.8 g / kg asphalt; 11.6 lb / ton asphalt for asphalt shingles; 13.0 g / kg asphalt; 26.0 lb / ton asphalt for roof tar). It should be noted that these are likely lower bound emissions as all roofing asphalt products are exposed to significant sunlight, which enhances emissions, and that emissions did not cease by the end of the experiment. For total roofing emissions, the prior emissions factors are weighted with the assumption that 5% of roofing asphalt is used for roof tar, with the remaining 95% predominantly being used for shingles (0.40 lb / ton asphalt for application; 12.32 lb / ton asphalt for in-use).

### Controls

There are no controls assumed for this category.

### Emissions

\begin{equation} 
  E_{c,a} = Usage_{c,a} \times \frac{EF_{a}}{2000}
  (\#eq:asphalt-emissions)
\end{equation} 

<div style="margin-left: 2em;">
  Where: <br> \newline
  \( E_{c,a} \) = Annual emissions in county *c* for application *a*, in short tons <br> \newline
  \( Usage_{c,a} \) = Liquid asphalt usage in county *c* for application *a*, in short tons <br> \newline
  \( EF_{a} \) = Emission factor for application *a*, in lb/ton asphalt <br> \newline
  \( a \) = Application types include hot-mix, warm-mix, cutback, and emulsified <br> \newline
</div>

### Sample Calculations

Table \@ref(tab:asphalt-sample-calculations) contains sample calculations for VOC emissions from emulsified asphalt (SCC: 2461022000). The values in these equations are demonstrating program logic and are not representative of any specific NEI year or county.

```{r, include=FALSE}
df <- data.frame(
  Eq = c(1:4),
  Equation = c("$Usage_{s,a} = Usage_{sp,a} \\times \\frac{HA_{s}}{HA_{sp}}$", 
               "Only applicable for hot- and warm-mix", 
               "$Usage_{c,a} = Usage_{s,a} \\times \\frac{\\sum_{r}PVMT_{c}}{\\sum_{r}PVMT_{s}}$", 
               "$E_{c,a} = Usage_{c,a} \\times \\frac{EF_{a}}{2000}$"),
  Values = c("$172 \\times \\frac{6.5}{19.9}$", 
             "--", 
             "$56 \\times \\frac{2.38E^9}{5.15E^{10}}$", 
             "$2.58 \\times \\frac{197.52}{2000}$"),
  Result = c("56 short tons of liquid asphalt usage for emulsified applications", 
             "--", 
             "2.58 short tons of liquid asphalt usage for emulsified applications", 
             "0.26 short tons of VOC emissions from emulsified asphalt")
)
```

```{r tidy=FALSE, echo=FALSE, booktabs = TRUE}
kableExtra::kbl(
  df, booktabs = TRUE, 
  escape = FALSE, # This is important to render LaTeX math properly
  caption = 'Sample Calculations',
  col.names = c("Eq. \\#", "Equation", "Values", "Result"),
  label = "asphalt-sample-calculations"
) %>%
kable_styling(full_width = FALSE, position = "center", latex_options = "scale_down")
```

### Improvements/Changes in the 2023 NEI

Small updates to the usage of asphalt cement for paving applications were made for the 2023 NEI. Specifically, since 2023-specific usage was not publicly released by the Asphalt Institute, other methods were required to scale 2020 values to 2023 (these are noted above). Methods for estimating emissions from roofing asphalts are new to the National Emissions Inventory process. Previously, select stakeholders submitted data for 2461023000. Starting with the 2023 NEI, EPA will generate default emissions estimates for this SCC throughout the nation.

### Puerto Rick and U.S. Virgin Islands

Insufficient data exists to calculate emissions for the counties in Puerto Rico and the US Virgin Islands. As such, emissions are based on two proxy counties in Florida: 12011 (Broward County) for Puerto Rico and 12087 (Monroe County) for the U.S. Virgin Islands. Per-capita emission factors from Broward County and Monroe County are applied to Puerto Rico and the U.S. Virgin Islands, respectively.

## References {#asphalt-references}

1. 2020 Asphalt Usage Survey for the United States and Canada. The Asphalt Institute, Lexington, KY.
2. Williams, B.A., J.R. Willis, & Shacat, J. (2021). Annual Asphalt Pavement Industry Survey on Recycled Materials and Warm-Mix Asphalt Usage: 2020, 11th Annual Survey (IS 138). National Asphalt Pavement Association, Greenbelt, Maryland. doi:10.13140/RG.2.2.14846.46409.
3. U.S. Department of Transportation, Federal Highway Administration, Center for Accelerating Innovation. [Warm Mix Asphalt FAQs](https://www.fhwa.dot.gov/innovation/everydaycounts/edc-1/wma-faqs.cfm#hot). 
4. Khare, P., Machesky, J., Soto, R., He, M., Presto, A.A., Gentner, D.R., Asphalt-related emissions are a major missing nontraditional source of secondary organic aerosol precursors. Science Advances, 6, 35, doi:10.1126/sciadv.abb9785.
5. U.S. Department of Transportation, Federal Highway Administration, Functional System Length, Table [HM-51 – Highways Statistics 2020](https://www.fhwa.dot.gov/policyinformation/statistics/2020/hm51.cfm).
6. [Annual Value of Private Nonresidential Construction Put in Place by State](https://www.census.gov/construction/c30/xlsx/nrstate.xlsx), 2008-2022, U.S. Census Bureau, 2023.
7. [Annual Value of State and Local Construction Put in Place by State, 2006-2022](https://www.census.gov/construction/c30/xlsx/slstate.xlsx), U.S. Census Bureau, 2023.
8. [County Population Totals: 2020-2022](https://www.census.gov/programs-surveys/popest/data/tables.html), U.S. Census Bureau, 2023.
9. Cutback Asphalt MSDS
```{r, include=FALSE}
df <- data.frame(
  Supplier = c("Valero", 
               "Asphalt Emulsion Industries", 
               "Martin Asphalt Company",
               "Mohawk Asphalt Emulsions",
               "Asphalt and Fuel Supply",
               "Valero",
               "Valero"),
  MSDS     = c("2013V04",
               "CUT-SDS-1",
               "Jan 2007",
               "UN1999",
               "211",
               "211",
               "210")
)
```

```{r tidy=FALSE, echo=FALSE, booktabs = TRUE}
kableExtra::kbl(
  df, booktabs = TRUE, 
  col.names = c("Product Supplier", "MSDS\\/SDS ID"),
  label = "cutback-MSDS"
) %>%
kable_styling(full_width = FALSE, position = "center", latex_options = c("scale_down", "H"))
```

```{r, include=FALSE}
df <- data.frame(
  Pollutant = c("Naphtha", 
               "Naphthalene & PAH\\*\\*\\*", 
               "Toluene\\*\\*\\*",
               "Xylene\\*\\*\\*",
               "Benzene\\*\\*\\*",
               "Ethylbenzene\\*\\*\\*",
               "Hydrogen Sulfide\\*\\*\\*", 
               "Total VOC\\*\\*"),
  Wght     = c("40",
                "0.58",
                "0.59",
                "0.99",
                "0.19",
                "0.49",
                "0.09",
                "--"),
  EmiFac    = c("760",
               "11.02",
               "11.21",
               "18.81",
               "3.61",
               "9.31",
               "1.71",
               "813.96")
)
```

```{r tidy=FALSE, echo=FALSE, booktabs = TRUE}
kableExtra::kbl(
  df, booktabs = TRUE, 
  col.names = c("Pollutant", "Avg. \\% by Weight", "Emission Factor [lb/ton]\\*"),
  label = "cutback-EF"
) %>%
kable_styling(full_width = FALSE, position = "center", latex_options = c("scale_down", "H"))
```
\noindent
\* = Assumes 95% volatilization <br> \newline
\*\* = Excludes hydrogen sulfide (not organic) <br> \newline
\*\*\* = Is a Hazardous Air Pollutant <br> \newline

10. Emulsified Asphalt MSDS
```{r, include=FALSE}
df <- data.frame(
  Supplier = c("Marathon", 
               "Marathon", 
               "Asphalt Emulsion Industries",
               "U.S. Oil and Refining Co."),
  MSDS     = c("0137MAR019",
               "0138MAR019",
               "EMU-SDS-1",
               "951")
)
```

```{r tidy=FALSE, echo=FALSE, booktabs = TRUE}
kableExtra::kbl(
  df, booktabs = TRUE, 
  col.names = c("Product Supplier", "MSDS\\/SDS ID"),
  label = "cutback-MSDS"
) %>%
kable_styling(full_width = FALSE, position = "center", latex_options = c("scale_down", "H"))
```

```{r, include=FALSE}
df <- data.frame(
  Pollutant = c("Naphtha", 
               "Naphthalene & PAH\\*\\*\\*", 
               "Hydrogen Sulfide\\*\\*\\*", 
               "Total VOC\\*\\*"),
  Wght     = c("10",
                "0.29",
                "0.09",
                "--"),
  EmiFac    = c("190",
               "5.51",
               "1.71",
               "195.51")
)
```

```{r tidy=FALSE, echo=FALSE, booktabs = TRUE}
kableExtra::kbl(
  df, booktabs = TRUE, 
  col.names = c("Pollutant", "Avg. \\% by Weight", "Emission Factor [lb/ton]\\*"),
  label = "cutback-EF"
) %>%
kable_styling(full_width = FALSE, position = "center", latex_options = c("scale_down", "H"))
```
\noindent
\* = Assumes 95% volatilization <br> \newline
\*\* = Excludes hydrogen sulfide (not organic) <br> \newline
\*\*\* = Is a Hazardous Air Pollutant <br> \newline