--- 
documentclass: book
classoption: oneside
biblio-style: apalike
link-citations: yes
colorlinks: yes
lot: yes
lof: yes
fontsize: 11pt
---

```{r, include=FALSE}
library(readxl)
library(knitr)
library(kableExtra)
library(readr)
library(ggplot2)
library(dplyr)
library(scales)
```

# Acronyms and Chemical Notations {-}

```{r, include=FALSE}
df <- read_excel("./tables/Section-0/acronyms.xlsx")
```

```{r table-single, tidy=FALSE, echo=FALSE}
knitr::kable(df, longtable = TRUE, booktabs = TRUE, col.names=NULL)
```

<!--chapter:end:00-acronyms.Rmd-->

--- 
documentclass: book
classoption: oneside
biblio-style: apalike
link-citations: yes
colorlinks: yes
lot: yes
lof: yes
fontsize: 11pt
---

```{r, include=FALSE}
library(readxl)
library(knitr)
library(kableExtra)
library(readr)
library(ggplot2)
library(dplyr)
library(scales)
library(glue)
```

# Cooking {#cooking}

## Sector Descriptions and Overview 

Cooking is a source of both gaseous and particulate pollutants [[ref 1](#cooking-references) – [ref 7](#cooking-references)]. Primary particulate emissions from cooking are predominantly organic carbon [[ref 2](#cooking-references), [ref 5](#cooking-references)] and positive matrix factorization analysis of Aerosol Mass Spectrometry data from multiple cities in the United States suggest cooking contributes ~16% of observed organic aerosol and ~8% of observed PM2.5 in urban areas [[ref 8](#cooking-references) – [ref 9](#cooking-references)]. In the NEI, emissions are estimated from the cooking of meat, including steak, hamburger, poultry, pork, and seafood, and french fries on five different cooking devices: chain-driven (conveyorized) charbroilers, underfired charbroilers, deep-fat fryers, flat griddles and clamshell griddles. The table below notes all SCCs covered in this source category and the SCCs for which the EPA generates default emissions. 

Table \@ref(tab:cooking-sccs) notes all SCCs covered in this source category and the SCCs for which the EPA generates default emissions. 

```{r, include=FALSE}
df <- read_excel("./tables/Section-19/cooking-sccs.xlsx")
df[is.na(df)] <- ""
```

```{r tidy=FALSE, echo=FALSE, booktabs = TRUE}
kableExtra::kbl(
  df, booktabs = TRUE,
  caption = 'Cooking SCCs in the 2023 NEI.',
  label = 'cooking-sccs'
) %>%
kable_styling(full_width = FALSE, position = "center", latex_options = "scale_down")
```

## EPA-developed estimates

To estimate emissions from commercial establishments, the amount of meat and french fries cooked on various cooking devices in each county is estimated. These estimates are year-specific and based on activity statistics from a 2001 telephone survey in southern California [[ref 10]](#cooking-references) and annually varying restaurant count statistics from Open Street Map [[ref 11]](#cooking-references). These meat consumption estimates are then scaled to match estimates of commercially consumed meat consumption statistics from the U.S. Department of Agriculture’s Food Availability Data System [[ref 12]](#cooking-references). The amount of french fries cooked by the foodservice industry is from a report prepared for Potatoes USA [[ref 13]](#cooking-references). For residential cooking sources, the mass of residentially consumed meat consumption statistics from the U.S. Department of Agriculture’s Food Availability Data System are also used. The total amount of meat or french fries cooked on each device is then multiplied by emissions factors for relevant pollutants to estimate emissions from cooking.

### Activity Data

Activity data is the amount of meat consumed away from home and at home. For commercial cooking, these estimates are based on a 2001 telephone survey in southern California [[ref 10]](#cooking-references) that reports the fraction of restaurants in a county that use different pieces of commercial cooking equipment by restaurant-type, the average number of cooking devices per restaurant, and the average mass of meat cooked on each device. The county-level number of various restaurant-types are retrieved from Open Street Map [[ref 11]](#cooking-references). Restaurant count statistics are pulled from this database using the SIC codes listed in Table \@ref(tab:sic-codes).

```{r, include=FALSE}
df <- read_excel("./tables/Section-19/sic-codes.xlsx")
df[is.na(df)] <- ""
```

```{r tidy=FALSE, echo=FALSE, booktabs = TRUE}
kableExtra::kbl(
  df, booktabs = TRUE,
  caption = 'SIC codes used to classify restaurant types.',
  label = 'sic-codes'
) %>%
kable_styling(full_width = FALSE, position = "center", latex_options = "scale_down")
```

County-level restaurant count statistics are then multiplied by the fraction of restaurants in a county that use different pieces of commercial cooking equipment by restaurant-type, the average number of cooking devices per restaurant, and the average mass of meat cooked on each device. These values are summarized in Tables \@ref(tab:fraction-appliance) - Tables \@ref(tab:mass-appliance).

```{r, include=FALSE}
df <- read_excel("./tables/Section-19/fraction-appliance.xlsx")
df[is.na(df)] <- ""
```

```{r tidy=FALSE, echo=FALSE, booktabs = TRUE}
kableExtra::kbl(
  df, booktabs = TRUE,
  caption = 'Fraction of restaurants in a county that use different pieces of commercial cooking equipment by restaurant-type.',
  label = 'fraction-appliance'
) %>%
kable_styling(full_width = FALSE, position = "center", latex_options = c("scale_down", "H"))
```

```{r, include=FALSE}
df <- read_excel("./tables/Section-19/number-appliance.xlsx")
df[is.na(df)] <- ""
```

```{r tidy=FALSE, echo=FALSE, booktabs = TRUE}
kableExtra::kbl(
  df, booktabs = TRUE,
  caption = 'Average number of cooking devices per restaurant. Average of non-specific charbroiler and griddle used, where necessary.',
  label = 'number-appliance'
) %>%
kable_styling(full_width = FALSE, position = "center", latex_options = c("scale_down", "H"))
```

```{r, include=FALSE}
df <- read_excel("./tables/Section-19/mass-appliance.xlsx")
df[is.na(df)] <- ""
last_col_index <- ncol(df)
df[[2]] <- round(as.numeric(df[[2]]), 1)
df[[last_col_index]] <- round(as.numeric(df[[last_col_index]]), 1)
```

```{r tidy=FALSE, echo=FALSE, booktabs = TRUE}
kableExtra::kbl(
  df, booktabs = TRUE,
  caption = 'Average mass of meat cooked per year on each cooking device (tons).',
  label = 'mass-appliance'
) %>%
kable_styling(full_width = FALSE, position = "center", latex_options = c("scale_down", "H"))
```

The total mass of each type of meat commercially cooked on each appliance-type in a county is then estimated using the following equation:

\begin{equation} 
  M{c,t,a} = \sum_{r}(R_{c,a} \times A_{r,a} \times N_{r,a}) \times T_{t,a} \times SF_{t}
  (\#eq:mass-meat)
\end{equation} 

<div style="margin-left: 2em;">
  Where: <br> \newline
  \( M_{c,t,a} \) = Estimated mass of commercially cooked meat in county (*c*) for each meat-type (*t*) on each appliance (*a*), in tons <br> \newline
  \( R_{c,r} \) = Number of restaurants in county (*c*) for each restaurant-type (*r*), in # <br> \newline
  \( A_{r,a} \) = Estimated percent of restaurants for each restaurant-type (*r*) containing > 1 appliance (*a*), in % <br> \newline
  \( N_{r,a} \) = Estimated number of appliances (*a*) for each restaurant-type (*r*), when appliance (*a*) is present, in # <br> \newline
  \( T_{t,a} \) = Average mass of meat cooked for each meat-type (*t*) on each appliance (*a*), in tons/appliance <br> \newline
  \( SF_{t} \) = Scale factor for each meat-type (*t*) applied to ensure national-level meat consumption matches USDA statistics, in % <br> \newline
  \( c \) = County <br> \newline
  \( t \) = Meat-Type (e.g., Steak, Hamburger, etc.) <br> \newline
  \( r \) = Restaurant-Type (e.g., Fast Food, Family, Seafood, etc.) <br> \newline
  \( a \) = Appliance (e.g., Conveyorized Charbroiler, Underfired Charbroiler, etc.) <br> \newline
</div>

The Scale Factor in Eqn. \@ref(eq:mass-meat) is year-specific and derived using meat consumption statistics from the U.S. Department of Agriculture’s Food Availability Data System [[ref 12]](#cooking-references). Data on the total retail consumption (disappearance) of beef, pork, veal, and lamb, chicken and turkey, and seafood are retrieved, and using data from the USDA’s Food Consumption and Nutrition Intakes, it is assumed that ~65% of meat is consumed at home and ~35% is consumed away from home (restaurants, fast food, school, etc.). These percentages are subsequently used to estimate consumption at home and scale commercial cooking meat consumption. In Table \@ref(tab:USDA-consumption), total residential and commercial meat consumption from the USDA’s Food Availability Data System by meat-type estimates are provided for recent years. Table \@ref(tab:scale-factor) illustrates the derivation of the Scale Factor used in Eqn. \@ref(eq:mass-meat) for 2021.

```{r, include=FALSE}
df <- read_excel("./tables/Section-19/USDA-consumption.xlsx")
df[is.na(df)] <- ""
last_col_index <- ncol(df)
df[[last_col_index]] <- as.numeric(df[[last_col_index]])
```

```{r tidy=FALSE, echo=FALSE, booktabs = TRUE}
kableExtra::kbl(
  df, booktabs = TRUE,
  caption = 'Total Retail Consumption from USDA Food Availability Data System',
  label = 'USDA-consumption'
) %>%
kable_styling(full_width = FALSE, position = "center", latex_options = c("scale_down", "H"))
```
\noindent
\* = Consumption statistics for fresh and frozen fish and shellfish is not available after 2018

```{r, include=FALSE}
df <- data.frame(
  Quantity = c("USDA Consumption at Home\\*", 
               "USDA Commercial Consumption\\*", 
               "Base Methods Commercial Consumption", 
               "Scale Factor"),
  Steak     = c("3,172,566",
                "1,708,305",
                "8,583,805",
                "0.20"),
  Hamburger = c("3,172,566",
                "1,708,305",
                "18,206,818",
                "0.09"),
  Poultry  = c("12,158,936",
               "6,547,120",
               "22,104,545",
               "0.30"),
  Pork     = c("5,457,592",
               "2,938,703",
               "3,986,449",
               "0.74"),
  Seafood  = c("1,305,599",
               "700,901",
               "8,392,852",
               "0.08"),
  Other    = c("163,281",
               "87,921",
               "8,888,702",
               "0.01")
)
```

```{r tidy=FALSE, echo=FALSE, booktabs = TRUE}
kableExtra::kbl(
  df, booktabs = TRUE, 
  caption = 'Scale Factor Derivation',
  col.names = c("Quantity", "Steak\\*\\*", "Hamburger\\*\\*", "Poultry", "Pork", "Seafood", "Other"),
  label = "scale-factor"
) %>%
kable_styling(full_width = FALSE, position = "center", latex_options = c("scale_down", "H"))
```
\noindent
\* = Using data from the USDA’s Food Consumption and Nutrition Intakes, it is assumed that ~65% of meat is consumed at home and ~35% is consumed away from home (restaurants, fast food, school, etc.). 
\*\* = It is assumed that half of all beef consumption reported in the USDA’s Food Availability Data System is steak and half is hamburger.

The amount of french fries cooked in each county is calculated based on the amount of frozen potatoes used in the foodservice industry. The total amount of french fries cooked is reported at the national level. The process used to distribute the national amount of french fries cooked to the county-level is discussed in the next section.

### Allocation Procedure



### Emission Factors



### Controls

There are no controls assumed for this category.

### Emissions

\begin{equation} 
  E_{c,a} = Usage_{c,a} \times \frac{EF_{a}}{2000}
  (\#eq:cooking-emissions)
\end{equation} 

<div style="margin-left: 2em;">
  Where: <br> \newline
  \( E_{c,a} \) = Annual emissions in county *c* for application *a*, in short tons <br> \newline
  \( Usage_{c,a} \) = Liquid asphalt usage in county *c* for application *a*, in short tons <br> \newline
  \( EF_{a} \) = Emission factor for application *a*, in lb/ton asphalt <br> \newline
  \( a \) = Application types include hot-mix, warm-mix, cutback, and emulsified <br> \newline
</div>

### Sample Calculations

Table \@ref(tab:cooking-sample-calculations) contains sample calculations for VOC emissions from emulsified asphalt (SCC: 2461022000). The values in these equations are demonstrating program logic and are not representative of any specific NEI year or county.

```{r, include=FALSE}
df <- data.frame(
  Eq = c(1:4),
  Equation = c("$Usage_{s,a} = Usage_{sp,a} \\times \\frac{HA_{s}}{HA_{sp}}$", 
               "Only applicable for hot- and warm-mix", 
               "$Usage_{c,a} = Usage_{s,a} \\times \\frac{\\sum_{r}PVMT_{c}}{\\sum_{r}PVMT_{s}}$", 
               "$E_{c,a} = Usage_{c,a} \\times \\frac{EF_{a}}{2000}$"),
  Values = c("$172 \\times \\frac{6.5}{19.9}$", 
             "--", 
             "$56 \\times \\frac{2.38E^9}{5.15E^{10}}$", 
             "$2.58 \\times \\frac{197.52}{2000}$"),
  Result = c("56 short tons of liquid asphalt usage for emulsified applications", 
             "--", 
             "2.58 short tons of liquid asphalt usage for emulsified applications", 
             "0.26 short tons of VOC emissions from emulsified asphalt")
)
```

```{r tidy=FALSE, echo=FALSE, booktabs = TRUE}
kableExtra::kbl(
  df, booktabs = TRUE, 
  escape = FALSE, # This is important to render LaTeX math properly
  caption = 'Sample Calculations',
  col.names = c("Eq. \\#", "Equation", "Values", "Result"),
  label = "cooking-sample-calculations"
) %>%
kable_styling(full_width = FALSE, position = "center", latex_options = "scale_down")
```

### Improvements/Changes in the 2023 NEI

-	Sector changed from “Commercial Cooking” to “Cooking” to include both commercial and residential cooking.
-	Scale Factor are applied to the average amount of meat cooked on each cooking device to ensure national-level meat consumption matches USDA statistics.
-	Emission Factors were updated


### Puerto Rick and U.S. Virgin Islands

Insufficient data exists to calculate emissions for the counties in Puerto Rico and the US Virgin Islands. As such, emissions are based on two proxy counties in Florida: 12011 (Broward County) for Puerto Rico and 12087 (Monroe County) for the U.S. Virgin Islands. Per-capita emission factors from Broward County and Monroe County are applied to Puerto Rico and the U.S. Virgin Islands, respectively.

## References {#cooking-references}

1.	R. Fortmann, P. Kariher and R. Clayton, Indoor Air Quality: Residential Cooking Exposures, ARCADIS Geraghty & Miller, Inc., Prepared for the State of California Air Resources Board; Contract Num: 97-330, 2001.
2.	N. Gysel, W. A. Welch, C. L. Chen, P. Dixit, D. R. Cocker, 3rd and G. Karavalakis, Particulate matter emissions and gaseous air toxic pollutants from commercial meat cooking operations, Journal of Environmental Science, 2018, 65, 162-170.
3.	F. Klein, S. M. Platt, N. J. Farren, A. Detournay, E. A. Bruns, C. Bozzetti, K. R. Daellenbach, D. Kilic, N. K. Kumar, S. M. Pieber, J. G. Slowik, B. Temime-Roussel, N. Marchand, J. F. Hamilton, U. Baltensperger, A. S. Prevot and I. El Haddad, Characterization of Gas-Phase Organics Using Proton Transfer Reaction Time-of-Flight Mass Spectrometry: Cooking Emissions, Environ Sci Technol, 2016, 50, 1243-1250.
4.	T. Liu, Z. Wang, D. D. Huang, X. Wang and C. K. Chan, Significant Production of Secondary Organic Aerosol from Emissions of Heated Cooking Oils, Environ Sci Tech Let, 2018, 5, 32-37.
5.	J. D. McDonald, B. Zielinska, E. M. Fujita, J. C. Sagebiel, J. C. Chow and J. G. Watson, Emissions from charbroiling and grilling of chicken and beef, J Air Waste Manag Assoc, 2003, 53, 185-194.
6.	J. J. Schauer, M. J. Kleeman, G. R. Cass and B. R. T. Simoneit, Measurement of Emissions from Air Pollution Sources. 1. C1 through C29 Organic Compounds from Meat Charbroiling, Environ Sci Technol, 1999, 33, 1566-1577.
7.	J. J. Schauer, M. J. Kleeman, G. R. Cass and B. R. T. Simoneit, Measurement of Emissions from Air Pollution Sources. 4. C1-C27 Organic Compounds from Cooking with Seed Oils, Environ Sci Technol, 2002, 36, 567-575.
8.	P. L. Hayes, A. M. Ortega, M. J. Cubison, K. D. Froyd, Y. Zhao, S. S. Cliff, W. W. Hu, D. W. Toohey, J. H. Flynn, B. L. Lefer, N. Grossberg, S. Alvarez, B. Rappenglueck, J. W. Taylor, J. D. Allan, J. S. Holloway, J. B. Gilman, W. C. Kuster, J. A. De Gouw, P. Massoli, X. Zhang, J. Liu, R. J. Weber, A. L. Corrigan, L. M. Russell, G. Isaacman, D. R. Worton, N. M. Kreisberg, A. H. Goldstein, R. Thalman, E. M. Waxman, R. Volkamer, Y. H. Lin, J. D. Surratt, T. E. Kleindienst, J. H. Offenberg, S. Dusanter, S. Griffith, P. S. Stevens, J. Brioude, W. M. Angevine and J. L. Jimenez, Organic aerosol composition and sources in Pasadena, California, during the 2010 CalNex campaign, J Geophys Res-Atmos, 2013, 118, 9233-9257.
9.	Y. L. Sun, Q. Zhang, J. J. Schwab, K. L. Demerjian, W. N. Chen, M. S. Bae, H. M. Hung, O. Hogrefe, B. Frank, O. V. Rattigan and Y. C. Lin, Characterization of the sources and processes of organic and inorganic aerosols in New York city with a high-resolution time-of-flight aerosol mass apectrometer, Atmos Chem Phys, 2011, 11, 1581-1602.
10.	M. Potepan, Charbroiling Activity Estimation, Public Research Institute, Prepared for the California Air Resources Board and the California Environmental Protection Agency; Contract Num: 98-721, 2001.
11. [Open Street Map](https://www.openstreetmap.org/), 2023.
12. United States Department of Agriculture, [Food Availability per Capita Data System](https://www.ers.usda.gov/data-products/food-availability-per-capita-data-system/).
13. Technomic, 2020. [Domestic Sales and U.S. Potato Utilization Report](https://potatoesusa.com/research-reports/category/market-insights/). Prepared for Potatoes USA.
14. Norbeck, Joseph, 1997. Further Development of Emission Test Methods and Development of Emission Factors for Various Commercial Cooking Operations. Prepared for the South Coast Air Quality Management District.
15.	C. A. Alves, M. Evtyugina, E. Vicente, A. Vicente, C. Goncalves, A. I. Neto, T. Nunes and N. Kovats, 2022. “Outdoor charcoal grilling: Particulate and gas-phase emissions, organic speciation and ecotoxicological assessment.” Atmos Environ. 285.

<!--chapter:end:19-cooking.Rmd-->

--- 
documentclass: book
classoption: oneside
biblio-style: apalike
link-citations: yes
colorlinks: yes
lot: yes
lof: yes
fontsize: 11pt
---

```{r, include=FALSE}
library(readxl)
library(knitr)
library(kableExtra)
library(readr)
library(ggplot2)
library(dplyr)
library(scales)
```

# Solvents: Asphalt {#asphalt}

## Sector Descriptions and Overview 

Liquid asphalt is a petroleum-derived substance used in paving applications, such as the construction of roads, parking lots, driveways, and airport runways, as well as non-paving applications, such as the manufacturing of roofing shingles. While liquid asphalt can be found in natural deposits, most is produced from crude oil. Vacuum distillation separates components of crude oil based on boiling point. Products generated from this process include naptha, gasoline, diesel, and liquid asphalt, the last of which has a boiling point greater than 500 °C. As a result, most volatile, light fractions of organics are separated from liquid asphalt during distillation and prior to use.

In paving applications, liquid asphalt can be applied cold or heated. If applied cold, additional components must be added to lower the viscosity of the material, which allows it to be spread upon a surface (e.g., roadway surface). Cutback asphalt (SCC: 2461021000) is a cold application process that involves mixing the liquid asphalt with petroleum solvents (e.g., naphtha, kerosene, fuel oil, diesel, etc.). Following application, these higher volatility solvents evaporate, leaving the asphalt in place. Due to this increased organic emissions potential, cutback asphalts have grown less common over time and this process now constitutes ~1% of liquid asphalt is use [[ref 1]](#ref-1). Emulsified asphalt (SCC: 2461022000) is a separate cold application process that utilizes water-based solvents and an emulsifying agent. The result is a stable liquid suspension with asphalt globules. Following application, the additives evaporate and leave the asphalt in place. In contrast to cutback asphalts, emulsified asphalts have become more common in recent years and ~10% of liquid asphalt is now used in these applications.

Liquid asphalt can also be applied heated, which both lowers the viscosity of the material and minimizes the need for added solvents. Hot-mix asphalt application (SCC: 2461025100), which is the traditional method for asphalt pavement production, involves combining the liquid asphalt with aggregate at a hot-mix plant and heating the mixture to +150 °C. The mixture is then hauled to the usage site heated, where it is placed, compacted, and ambiently cooled. This process does not require any solvent additions. Warm-mix asphalt application (SCC: 2461025200) is a more recent technology that enables asphalt pavement production to occur at 20 – 40 °C cooler temperatures than hot-mix asphalt application. Warm-mix asphalt applications at reduced temperatures now constitute ~18% of asphalt paving applications [[ref 2]](#ref-2). These lower production temperatures promote energy savings through reductions in fuel use and lower emissions at the hot-mix plant. To reduce the viscosity of the liquid asphalt in warm-mix applications, water, water-bearing minerals, chemicals, waxes, organic additives, or a combination of technologies must be added [[ref 3]](#ref-3). 

While heated applications processes represent most liquid asphalt paving applications, it has historically been assumed that emissions were sparse due to the removal of the more volatile organics during distillation. Recent research has demonstrated that less volatile organic vapors from liquid asphalt do evaporate at temperatures associated with hot-mix application (~140 °C), warm-mix application (120 °C), and post-application, or “in-use” temperatures [[ref 4]](#ref-4). Emission during the “in-use” period diffuse from the pavement over time following application. 

Roofing asphalts include asphalt cements and emulsions used in the manufacturing of asphalt shingles, asphalt sealant, and roof tar. In 2020, manufacturing of these products consumed ~1.95 million short tons of asphalt, which is ~9% of all asphalt generated in the United States. These materials are predominantly applied in ambient conditions (i.e., as asphalt shingles). Roof tar, which is applied hot and at temperatures like hot-mix road asphalt (+150 °C), comprises ~5% of roofing asphalt usage. In addition, all forms of roofing asphalt are exposed to both high temperatures and solar radiation throughout the duration of their life cycles, which have the potential of generating enhancements of emissions.

The table below notes all SCCs covered in this source category and the SCCs for which the EPA generates default emissions. 

```{r, include=FALSE}
df <- read_excel("./tables/Section-1/point-thresholds.xlsx")
```

```{r point-thres, tidy=FALSE, echo=FALSE}
knitr::kable(
  df, booktabs = TRUE,
  caption = 'Point source reporting thresholds, as potential to emit, for CAPs in the AERR. Note that the potential to emit is shown in tons per year, as defined in 40 CFR part 70, with the exception of lead.',
)
```

## EPA-developed estimates

Usage of liquid asphalt at the state-level for each process is calculated and subsequently allocated to the county-level using estimated vehicle miles traveled on paved roads and construction expenditure statistics for roofing processes. Emission factors consider both application and in-use processes. Net county-level emissions are quantified by multiplying the SCC-specific liquid asphalt usage by SCC-specific emission factors. The sources of data, calculation of state-level, SCC-specific usage results, allocation of state-level usage to the county-level, emission factors, and emission estimates are all discussed in subsequent sections.

### Activity Data

Activity data for these sources are the amount of liquid asphalt used in each process. This includes cutback, emulsified, hot mix, warm mix, and roofing. Each year, the Asphalt Institute releases an asphalt usage survey for the Unites States and Canada that reflects usage among their membership [[ref 1]](#ref-1). The Asphalt Institute estimates that their membership captures 90% of the United States market and the survey includes usage for cutback, emulsified, a summation of heated application processes, and roofing asphalt at the level of Petroleum Administration for Defense Districts (PADD) and sub-PADD levels. PADDs are geographic aggregations of the 50 states and the District of Columbia that were generated during World War II for purposes of administering oil allocation. At the sub-PADD level, usage resolution is provided in aggregations that include up to six states. Following 2020, the Asphalt Institute no longer publicly released this annual survey. As such, additional methods were used to scale the 2020 Asphalt Institute survey to 2023 values, which are further described below.

Separately, state-level, year-specific data on the production of asphalt pavement (i.e., mass of asphalt binder and aggregate) for heated application and the proportions used in warm mix processes at reduced temperatures are available from a National Asphalt Pavement Association (NAPA) annual report [[ref 1]](#ref-1). This report is not used in-lieu of the Asphalt Institute report because it does not include usage of cutback, emulsified, and roofing asphalts. However, the state-level heated application usage and reduced temperature (i.e., warm mix usage) proportions are used to allocate cutback, emulsified, hot mix, and warm mix usage from the Asphalt Institute survey to the state-level. Therefore, it is assumed that the state-level proportions of liquid paving asphalt used in heated applications within a sub-PADD match the state-level proportions of liquid asphalt used in cold application (i.e., cutback and emulsified) within a sub-PADD. Furthermore, since 2023 data is available from the NAPA annual report, the state-level growth or decline in asphalt pavement usage between 2020 and 2023 from these reports are used to scale the 2020 Asphalt Institute survey to generate 2023 estimates of all paving processes. Allocation of activity data for roofing asphalt utilizes separate methods and is described below. The derivation of state-level, per-application usage is as follows:

\begin{equation} 
  Usage_{s,a} = Usage_{sp,a} \times \frac{HA_{s}}{HA_{sp}}
  (\#eq:state-usage)
\end{equation} 

<div style="margin-left: 2em;">
  Where: <br> \newline
  \( Usage_{s,a} \) = Liquid asphalt usage in state *s* for application *a*, in short tons <br> \newline
  \( Usage_{sp,a} \) = sub-PADD usage of liquid asphalt in sub-PADD *sp* associated with state *s* for application *a* from the Asphalt Institute survey, in short tons <br> \newline
  \( HA_{s} \) = Heated application usage in state *s* from the NAPA survey, in short tons <br> \newline
  \( HA_{sp} \) = Heated application usage in sub-PADD *sp* associated with state *s* from the NAPA survey, in short tons <br> \newline
  \( a \) = Application types include hot-mix, warm-mix, cutback, and emulsified <br> \newline
  \( sp \) = sub-PADD associated with state *s*. sub-PADDs include the 11 districts included in the Asphalt Institute survey [[ref 1]](#ref-1) <br> \newline
</div>

An additional transformation must be done to split the state-level, heated application usage into hot-mix and warm-mix application, respectively.

\begin{equation} 
\[ Usage_{s,a} = 
  \begin{aligned}
  &\begin{cases} 
    Usage_{s} \times \frac{WMA_{s}}{HA_{s}} & \text{for warm-mix application}\\ 
    Usage_{s} \times \frac{(1 - WMA_{s})}{HA_{s}} & \text{for hot-mix application }
  \end{cases}
  \end{aligned}
  (\#eq:piecewise-usage)
\]
\end{equation}

<div style="margin-left: 2em;">
  Where: <br> \newline
  \( Usage_{s,a} \) = Liquid asphalt usage in state *s* for application *a*, in short tons <br> \newline
  \( Usage_{s} \) = Liquid asphalt usage in state *s* for heated application, as derived in \@ref(eq:state-usage), in short tons <br> \newline
  \( WMA_{s} \) = Warm-mix application usage at reduced temperatures in state *s* from the NAPA survey, in short tons <br> \newline
  \( HA_{s} \) = Heated application usage in state *s* from the NAPA survey, in short tons. <br> \newline
  \( a \) = Application types include hot-mix and warm-mix. State-level summation across application types in \@ref(eq:piecewise-usage) yield the state-level application usage for heated application from \@ref(eq:state-usage). <br> \newline
</div>

### Allocation Procedure

### Emission Factors

### Controls

There are no controls assumed for this category.

### Emissions

### Sample Calculations

### Improvements/Changes in the 2023 NEI

Small updates to the usage of asphalt cement for paving applications were made for the 2023 NEI. Specifically, since 2023-specific usage was not publicly released by the Asphalt Institute, other methods were required to scale 2020 values to 2023 (these are noted above). Methods for estimating emissions from roofing asphalts are new to the National Emissions Inventory process. Previously, select stakeholders submitted data for 2461023000. Starting with the 2023 NEI, EPA will generate default emissions estimates for this SCC throughout the nation.

### Puerto Rick and U.S. Virgin Islands

Insufficient data exists to calculate emissions for the counties in Puerto Rico and the US Virgin Islands. As such, emissions are based on two proxy counties in Florida: 12011 (Broward County) for Puerto Rico and 12087 (Monroe County) for the U.S. Virgin Islands. Per-capita emission factors from Broward County and Monroe County are applied to Puerto Rico and the U.S. Virgin Islands, respectively.

## References

1. <a id="ref-1"></a> 2020 Asphalt Usage Survey for the United States and Canada. The Asphalt Institute, Lexington, KY.
2. <a id="ref-2"></a> Williams, B.A., J.R. Willis, & Shacat, J. (2021). Annual Asphalt Pavement Industry Survey on Recycled Materials and Warm-Mix Asphalt Usage: 2020, 11th Annual Survey (IS 138). National Asphalt Pavement Association, Greenbelt, Maryland. doi:10.13140/RG.2.2.14846.46409.
3. <a id="ref-3"></a> U.S. Department of Transportation, Federal Highway Administration, Center for Accelerating Innovation. [Warm Mix Asphalt FAQs](https://www.fhwa.dot.gov/innovation/everydaycounts/edc-1/wma-faqs.cfm#hot).
4. <a id="ref-4"></a> Khare, P., Machesky, J., Soto, R., He, M., Presto, A.A., Gentner, D.R., Asphalt-related emissions are a major missing nontraditional source of secondary organic aerosol precursors. Science Advances, 6, 35, 2020. doi:10.1126/sciadv.abb9785.
5. <a id="ref-5"></a> U.S. Department of Transportation, Federal Highway Administration, Functional System Length, [Table HM-51 – Highways Statistics 2020](https://www.fhwa.dot.gov/policyinformation/statistics/2020/hm51.cfm).

<!--chapter:end:31-asphalt.Rmd-->

